\documentclass[psamsfonts]{amsart}
%
%-------Packages---------
%
\usepackage[h margin=1 in, v margin=1 in]{geometry}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{tikz-cd}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{amsthm}
\usepackage{mathpazo}
\usepackage{yfonts}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage{fourier-orns}
\usepackage[all]{xy}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{url}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{pdfsync}
\usepackage{mathdots}
\usepackage{calligra}
%
\usepackage{tgpagella}
\usepackage[T1]{fontenc}
%
\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Matlab,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
  }
%
%--------Theorem Environments--------
%
\newtheorem{thm}{Theorem}[section]
\newtheorem*{thm*}{Theorem}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem*{lem*}{Lemma}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{quest}[thm]{Question}
%
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{defn*}{Definition}
\newtheorem{defns}[thm]{Definitions}
\newtheorem{con}[thm]{Construction}
\newtheorem{exmp}[thm]{Example}
\newtheorem{exmps}[thm]{Examples}
\newtheorem{notn}[thm]{Notation}
\newtheorem{notns}[thm]{Notations}
\newtheorem{addm}[thm]{Addendum}
\newtheorem{exer}[thm]{Exercise}
%
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem*{claim}{Claim}
\newtheorem*{aside*}{Aside}
\newtheorem*{rem*}{Remark}
\newtheorem*{hint*}{Hint}
\newtheorem*{note}{Note}
\newtheorem{rems}[thm]{Remarks}
\newtheorem{warn}[thm]{Warning}
\newtheorem{sch}[thm]{Scholium}
%
%--------Macros--------
\renewcommand{\qedsymbol}{$\blacksquare$}
\renewcommand{\hom}{\mathsf{Hom}}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\O}{\mathscr{O}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\ib}[1]{\textbf{\textit{#1}}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\A}{\mathbb{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\V}{\vec{v}}
\newcommand{\RP}{\mathbb{RP}}
\newcommand{\CP}{\mathbb{CP}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\GL}{\mathsf{GL}}
\newcommand{\SL}{\mathsf{SL}}
\newcommand{\SP}{\mathsf{SP}}
\newcommand{\SO}{\mathsf{SO}}
\newcommand{\SU}{\mathsf{SU}}
\newcommand{\gl}{\mathfrak{gl}}
\newcommand{\g}{\mathfrak{g}}
\newcommand{\inv}{^{-1}}
\newcommand{\bra}[2]{ \left[ #1, #2 \right] }
\newcommand{\ind}{\lambda \in \Lambda}
\newcommand{\set}[1]{\left\lbrace #1 \right\rbrace}
\newcommand{\abs}[1]{\left\lvert#1\right\rvert}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\transv}{\mathrel{\text{\tpitchfork}}}
\newcommand{\enumbreak}{\ \\ \vspace{-\baselineskip}}
\let\oldexists\exists
\renewcommand\exists{\oldexists~}
\let\oldL\L
\renewcommand\L{\mathfrak{L}}
\makeatletter
\newcommand{\tpitchfork}{%
  \vbox{
    \baselineskip\z@skip
    \lineskip-.52ex
    \lineskiplimit\maxdimen
    \m@th
    \ialign{##\crcr\hidewidth\smash{$-$}\hidewidth\crcr$\pitchfork$\crcr}
  }%
}
\makeatother
\newcommand{\bd}{\partial}
\newcommand{\lang}{\begin{picture}(5,7)
\put(1.1,2.5){\rotatebox{45}{\line(1,0){6.0}}}
\put(1.1,2.5){\rotatebox{315}{\line(1,0){6.0}}}
\end{picture}}
\newcommand{\rang}{\begin{picture}(5,7)
\put(.1,2.5){\rotatebox{135}{\line(1,0){6.0}}}
\put(.1,2.5){\rotatebox{225}{\line(1,0){6.0}}}
\end{picture}}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\im}{Im}
\DeclareMathOperator{\grap}{graph}
\DeclareMathOperator{\codim}{codim}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\inter}{Int}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\indx}{ind}
\DeclareMathOperator{\alt}{Alt}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\trace}{trace}
\DeclareMathOperator{\ad}{ad}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\Ad}{Ad}
\DeclareMathOperator{\Lie}{Lie}
\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\dv}{div}
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\sheafhom}{\mathscr{H}\text{\kern -3pt {\calligra\large om}}\,}
\newcommand*\myhrulefill{%
   \leavevmode\leaders\hrule depth-2pt height 2.4pt\hfill\kern0pt}
\newcommand\niceending[1]{%
  \begin{center}%
    \LARGE \myhrulefill \hspace{0.2cm} #1 \hspace{0.2cm} \myhrulefill%
  \end{center}}
\newcommand*\sectionend{\niceending{\decofourleft\decofourright}}
\newcommand*\subsectionend{\niceending{\decosix}}
\def\upint{\mathchoice%
    {\mkern13mu\overline{\vphantom{\intop}\mkern7mu}\mkern-20mu}%
    {\mkern7mu\overline{\vphantom{\intop}\mkern7mu}\mkern-14mu}%
    {\mkern7mu\overline{\vphantom{\intop}\mkern7mu}\mkern-14mu}%
    {\mkern7mu\overline{\vphantom{\intop}\mkern7mu}\mkern-14mu}%
  \int}
\def\lowint{\mkern3mu\underline{\vphantom{\intop}\mkern7mu}\mkern-10mu\int}
%
%--------Hypersetup--------
%
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=blue,
    urlcolor=black
}
%
%--------Solution--------
%
\newenvironment{solution}
  {\begin{proof}[Solution]}
  {\end{proof}}
%
%--------Graphics--------
%
%\graphicspath{ {images/} }

\begin{document}
\author{Jeffrey Jiang}
\title{The Pontryagin-Thom Theorem}
\maketitle
\setcounter{section}{1}
\section*{Framed Bordism}
%
Let $M$ be a smooth manifold of dimension $n$, and $Y \subset M$ a submanifold of codimension $q$, i.e. $\dim Y = n - q$. Then the tangent bundle $TY$ is naturally a subbundle of the ambient tangent bundle $TM\vert_Y$ over $Y$. The normal bundle of $Y$ in $M$, denoted $NY$, is the vector bundle over $Y$ with each fiber being the quotient space $T_pM / T_pY$. This gives us the short exact sequence
$$\begin{tikzcd}
0 \ar[r] & TY \ar[r] & TM\vert_Y \ar[r] & NY \ar[r] & 0
\end{tikzcd}$$
You can think of $NY$ as the vector bundle of normal vectors pointing outward from $Y$. 
%
\begin{defn}
A \ib{framing} of a codimension $q$ submanifold $Y \subset M$ is an isomorphism $NY \to R^q \times Y$. A \ib{framed submanifold} of a manifold $M$ is the data of a submanifold $Y \subset M$ and a framing $NY \to R^q \times Y$. We note that the isomorphism $NY \to R^q \times Y$ is equivalent data to a global frame (basis of sections) for $NY$, hence the name.
\end{defn}
For example, if we consider $S^1 \subset \R^2$, we have that the tangent space $T_pS^1 \cong p^\perp$. Therefore, an example of a framing for $S^1$ is the assignment $p \mapsto p$, since $p$ spans the orthogonal complement of $T_pS^1$, giving a global frame for $NS^1$.
%
\begin{defn}
Let $X_0$ and $X_1$ be framed codimension $q$ submanifolds of $M$. Then $X_1$ and $X_2$ are \ib{framed bordant} if there exists a framed submanifold $Y \subset [0,1] \times M$ such that
\begin{enumerate}
\item $Y \cap \set{i} \times M = \set{i} \times X_i$ for $i = 0,1$.
\item The framings at $Y \cap \set{i}$ agree with the framings of $X_i$.
\end{enumerate}
We denote the framed cobordism group of codimension $q$ as $\Omega^{\text{fr}}_{m-q}(M)$. 
\end{defn}
There's a little subtlety by what we mean by ``the framings agree." By this, we note that the boundary of $[0,1] \times M$ contains two copies of $M$ at the boundary, $\bd([0,1] \times M) = \set{0} \times M \coprod \set{1} \times M$, and the inclusion $M \hookrightarrow \set{i} \times M$ gives us an isomorphism of vector bundles $TM \cong T(\set{i} \times M) \subset T([0,1] \times M)$. Likewise, we get a isomorphism of normal bundles $NX_i \cong N(\set{i} \times Y)$, where the normal bundle of $\set{i} \times Y$ is taken inside of $\set{i} \times M$. Hence a framing of $\set{i} \times X_i$ is equivalent to a framing of $X_i$, and we ask for the framings to agree with the framing for $Y$. A trivial, but easy to visualize example is to take $S^1 \subset \R^2$, with the framing we gave earlier. Then the cylinder $[0,1] \times S^1 \subset [0,1] \times \R^2$ is a framed bordism $S^1 \to S^1$, where the framing is given by our previous framing at each $\set{t} \times S^1$.

In particular, there is one easy way to obtain a huge family of framed submanifolds. Let $M$ be an $m$ dimensional manifold, and $N$ an $n$ dimensional manifold with $n \leq m$. Then by Sards theorem, for any map $f : M \to N$, the set of critical points of $f$ forms a set of measure $0$. In other words, for almost any point $p \in N$, we have that $df_x : T_xM \to T_pN$ is a surjective map, giving us that $f\inv(p)$ is a submanifold of $M$ of codimension $n$. The tangent space $T_xf\inv(p)$ is naturally identified with the kernel of $df_x$, so $df_x$ factors through the quotient $T_xM / T_xf\inv(p)$, giving an isomorphism $T_xM / T_xf\inv(p) \to T_pN$. Therefore, fixing a basis for $T_pN$ gives us an isomorphism $T_xM / T_xf\inv(p) \to \R^n \times f\inv(p)$, which is the definition of a framing for $f\inv(p)$.
%
\setcounter{section}{3}
%
\setcounter{thm}{0}
%
\section*{The Pontryagin-Thom Theorem}
%
We now consider the special case where we have $M$ mapping into a sphere $S^q$. Let $[M, S^q]$ denote the set of all homotopy classes of maps $M \to S^q$. Fix a framed submanifold $Y \subset M$. Then it admits a \ib{tubular neighborhood}, which is the data of an open set $U \subset M$ containing $Y$ equipped with a submersion $U \to Y$, and a diffeomorphism $NY \to U$ such that 
$$\begin{tikzcd} 
NY \ar[rr] \ar[dr] && U \ar[dl] \\
& Y
\end{tikzcd}$$
commutes. You can think of $U$ as a ``fattened up" version of $Y$, and the submersion $U \to Y$ squishes $U$ onto $Y$. Then the framing for $Y$ gives us an isomorphism $NY \to \R^q \times Y$, and then taking the inverse diffeomorphism $U \to NY$, we obtain a map $U \to NY \to \R^q \times Y \to \R^q$, where the last map is projection. We denote the map as $h : U \to \R^q$.  Intuitively, the map $h(p)$ is the distance from $p$ to $Y$. Then by fixing a cutoff function $\rho$, we define the \ib{Pontryagin-Thom collapse map} of $Y$, $f_Y : M \to S^q$ by
$$f_Y(p) = \begin{cases} 
\frac{h(p)}{\rho(\abs{h(p)})} & p \in U \\
\infty & p \in M - U
\end{cases}$$
Where we identify $S^q$ as the one point compactification $\R^q \cup \set{\infty}$ via stereographic projection. Intuitively, the map collapses $Y$ to a point, (namely $0$), and smoothly stretches out $U$ onto the sphere before collapsing the rest of $M$ onto the point at infinity.
%
\begin{thm}
Let $\varphi : [M,S^q] \to \Omega^{\text{fr}}_{n - q}$ be the map taking $[f]$ to the preimage $f\inv(p)$ of a regular value, and let $\psi : \Omega^{\text{fr}}_{n-q} \to [M, S^q]$ be the map taking the bordism class $[Y]$ to the homotopy class of the collapse map $f_Y$. Then $\varphi$ and $\psi$ are inverse maps.
\end{thm}
\begin{proof}
There's a lot of details to check here, and not much time, so we'll list what needs to be shown, and the tools used to show them
\begin{enumerate}
\item Show $\varphi$ is independent of choice of $p$, this is seen by noting that we can compose with a rotation to move any point of $S^q$ onto our choice of $p$.
\item Show $\varphi$ is well defined on homotopy classes. This is shown by taking a homotopy $F$ from $f_0$ to $f_1$ and (after some fiddling with transversality) taking the framed submanifold $F\inv(p) \subset [0,1] \times M$, which will be a bordism from $f_0\inv(p)$ to $f_1\inv(p)$.
\item Show that $\psi$ is independent of cutoff function and tubular neighborhood. This amounts to retracting one neighborhood onto another, and showing a similar cutoff function would produce a homotopic map
\item Show that $\psi$ is well defined on boridsm classes. Given bordant submanifolds $Y_0$ and $Y_1$, let $X \subset [0,1]$ be the bordism between them. Then the collapse map $f_X$ is the homotopy from $f_{Y_0}$ to $f_{Y_1}$.
\item Show that $\varphi \circ \psi$ is the identity map. This is easily seen, since the collapse map for a submanifold has regular value $0$, and preimage $Y$ by construction
\item Show that $\psi \circ \varphi$ is identity. This one is a bit tricky. The key fact here is that given a map $f_0$, $Y = f_0\inv(p)$, then we have $df_0\vert_Y = df_Y\vert_Y$, which is enough to deduce they are homotopic with some work.
\end{enumerate}
\iffalse
We first need to show that $\varphi$ and $\psi$ are well-defined on equivalence classes. For $\varphi$, suppose we have two maps $f_0$ and $f_1$ that are homotopic via $F : M \times [0,1] \to S^q$. In the case that $p$ is not a regular value of $f_0$ or $f_1$, we can perturb either slightly to obtain a homotopic map that is. We can also perturb $F$ to obtain a map $F'$ that it is transverse to $\set{p}$ and such that it agrees with $F$ in a small neighborhood of $\set{0,1} \times M$, so $F'_0 = f_0$ and $F'_1 = f_1$. Then $F^{'-1}(p)$ is a submanifold of $M \times [0,1]$, and by construction, it is a framed bordism $f_0\inv(p) \to f_1\inv(p)$. In addition, it is independent of our choice of $p$, since we can compose with a rotation on $S^q$ that takes any point to our regular value.

To show that $\psi$ is well defined, we need to show that for submanifolds $Y_0$ and $Y_1$, that the collapse maps $f_{Y_0}$ and $f_{Y_1}$ are homotopic. In addition, we would need to show that it is independent of our choice of cutoff function and tubular neighborhood, but we'll hand wave those details away for now. Let $X \subset [0,1] \times M$ be a bordism $Y_0 \to Y_1$. Then taking a tubular neighborhood $U$ of $X$ gives us a collapse map $f_X : [0,1] \times M \to S^q$. We then note that this is a homotopy from $f_{Y_0}$ to $f_{Y_1}$, so $\psi$ is well-defined as well.

We note that $\varphi \circ \psi$ is identity, since given a framed submanifold $Y$, we have that $0$ is a regular value of $f_Y$ with preimage $Y$ by construction. For the other direction, given $f_0$, this determines a submanifold $f_0\inv(p)$, which in turn defines a collapse map $f_1$ for $f_0\inv(p)$,. We need to show that this collapse map is homotopic to $f_0$. As it turns out, our construction gives us that $df_0\vert_Y = df_1\vert_Y$, which is enough to deduce that $f_0 \simeq f_1$, after some fiddling.
\fi
\end{proof}
The theorem gives us a remarkable dictionary from bordism to algebraic topology. In the case that $M = S^n$, we have that the maps are actually group isomorphisms $\pi_n(S^q) \cong \Omega^{\text{fr}}_{n-q}(S^n)$. Going a bit further, if we had a bit more machinery we could establish an isomorphism from the ring of stable homotopy groups of the sphere, and the framed bordism ring.
%
\end{document}