\documentclass[abstract=on,twoside]{scrreprt}

\usepackage[MyBox]{fancychap}
\usepackage{utmathjournal}

\begin{document}
\chapter{Principal Bundles} % your article title goes here
\thispagestyle{empty}
\vspace{-2cm}
\begin{flushright}
	\textbf{Jeffrey Jiang}\\ % your name goes here
	\texttt{jeffjiang@utexas.edu} % your email address goes here
\end{flushright}

\section*{Why Fiber Bundles?}

Suppose we want to study some space $M$, which for our purposes is a smooth manifold.
One way to study $M$ is to study functions $M \to F$ for some fixed target space $F$,
e.g. a manifold , $\R$, or a vector space. This is a perfectly good method
of studying $M$, but is sometimes not enough. More often then not, we want to
study ``function" from $M$ into a vector space that varies over the base
manifold $M$. For example, a vector field $X \in \mathfrak{X}(M)$ is not really
a map $M \to \R^n$, it is an assignment to each $p \in M$ a vector in $T_pM$.
In this way, we are led to the study of a smoothly parameterized family of
vector spaces -- the tangent bundle $TM$. This leads us to define a fiber bundle.
%
\begin{definition}
Let $M$ and $F$ be a smooth manifolds. A \textbf{\textit{fiber bundle}} over $M$ with
model fiber $F$ is the data of a smooth manifold $E$ with a smooth surjective map
$\pi : E \to M$ such that for each $p \in M$, there exists an open set $U$ and a
diffeomorphism $\varphi : \pi^{-1}(U) \to U \times F$ such that
\[\begin{tikzcd}
\pi^{-1}(U) \ar[dr, "\pi"']\ar[rr, "\varphi"] && U \times F \ar[dl, "p_1"] \\
& U
\end{tikzcd}\]
where $p_1 : U \times F \to U$ is projection onto the first factor. The map
$\varphi : \pi^{-1} \to U \times F$ is called a \textbf{\textit{local trivialization}}.
The space $E$ is called the \textbf{\textit{total space}}, while the manifold $M$
is called the \textbf{\textit{base space}}. We often omit naming the map $\pi$
and denote the fiber $\pi^{-1}(x)$ by $E_x$.
\end{definition}
%
This definition captures the notion of a family of manifolds diffeomorphic to $F$ that
are smoothly parameterized by the base space $M$. We also have a notion of a morphism
between bundles.
%
\begin{definition}
Let $\pi_E E \to M$ and $\pi_F : F \to M$ be fiber bundles with model fiber $X$. A
\textbf{\textit{bundle homomorphism}} is the data of a smooth map $\varphi : E \to F$
such that the diagram
\[\begin{tikzcd}
E \ar[dr, "\pi_E"']\ar[rr, "\varphi"] && F \ar[dl, "\pi_F"]\\
& M
\end{tikzcd}\]
commutes.
\end{definition}

Our original motivation for thinking about fiber bundles was for a generalized notion of
a function. To this end, we specify a special class of maps associated to a fiber
bundle.
%
\begin{definition}
Let $\pi : E \to M$ be a a fiber bundle with model fiber $F$.
A \textbf{\textit{local section}} of $\pi : E \to M$ is a smooth map $\sigma : U \to E$
such that $\pi \circ \sigma = \id_U$ for some open set $U \subset M$. If $U = M$, we
call $\sigma $ a \textbf{\textit{global section}}. Equivalently, it is a smooth assignment
to each $p \in U$ a point in the fiber $E_p$. We denote the space of sections over
an open set $U$ as $\Gamma_U(E)$.
\end{definition}
%
A section of $E \to M$ can be thought of as our desired generalization of a function.
A map $M \to F$ is the same data as a section of the trivial bundle $M \times F \to M$.
However, not every fiber bundle is trivial -- there can be a nontrivial ``twisting."
An example of this is the M\"obius band. Can you see why this bundle over $S^1$ is not
isomorphic to the trivial bundle $S^1 \times [0,1]$? \\

We are especially interested in two special classes of fiber bundles that carry
additional structure -- the fibers of a vector bundle carry the extra structure of a
vector space, and the fibers of a principal bundle have the extra structure of a
$G$-torsor for a Lie group $G$.
%
\begin{definition}
A \textbf{\textit{vector bundle of rank}} $k$ is a fiber bundle $E \to M$ such that each
fiber $E_x$ has the structure of a $k$-dimensional vector space (usually over $\R$ or
$\C$). A \textbf{\textit{vector bundle homomorphism}} is a bundle homomorphism that
restricts to a linear map on each fiber.
\end{definition}
%
Vector bundles form a familiar family of fiber bundles, as tangent bundles,
cotangent bundles, and their associated tensor bundles are all vector bundles.
%
\begin{definition}
Let $G$ be a Lie group. a (left) \textbf{\textit{$G$-torsor}} is a smooth manifold $M$
equipped with a smooth (left) $G$-action that is free in transitive i.e.
\begin{enumerate}
	\item For any $g \in G$ and $p \in M$, if $g\cdot p = p$, then $g = e$.
	\item For any $p \in M$, the orbit $G \cdot p$ is all of $M$
\end{enumerate}
combined, these conditions give that for any fixed point $p$, there exists a unique
group element $g \in G$ that takes $p$ to any other point of $M$. In other words,
the map $g \mapsto g \cdot p$ is a diffeomorphism.
\end{definition}
%
\begin{example}
Let $V$ be a vector space. A \textbf{\textit{basis}} (also called a
\textbf{\textit{frame}}) of $V$ is a linear isomorphism $\R^n \to V$, and the
set of frames of $V$ is denoted $\mathcal{B}(V)$. There is a
natural right action of $GL_n\R$ on $\mathcal{B}(V)$ by precomposition, i.e.
for $g \in GL_n\R$ and $b \in \mathcal{B}(V)$, we have $b \cdot g = b \circ g$.
This action is free and transitive, which allows us to define a topology and
smooth structure on $\mathcal{B}(V)$, giving it the structure of a
right $GL_n\R$-torsor.
\end{example}
%
$G$-torsors can be thought of as Lie groups without a fixed choice of identity element.
Indeed, if we fix a point $p$ of a $G$-torsor $M$, this determines a unique Lie group
structure on $M$ where $p$ is the identity element via the diffeomorphism
$g \mapsto g \cdot p$. For example, there is no distinguished basis for a vector
space $V$, which is reflected by the fact that $\mathcal{B}(V)$ is a torsor, rather
than a group. This then allows us to define a principal $G$-bundle.
%
\begin{definition}
A \textbf{\textit{principal $G$-bundle}} is a fiber bundle $\pi : P \to M$ with a
smooth \emph{right} $G$-action such that every fiber $\pi^{-1}(x)$ is a $G$-torsor,
and every point $p \in M$ in the base has a local trivialization
$\pi^{-1}(U) \to U \times G$ that is $G$-equivariant, where the $G$-action on
$U \times G$ is right multiplication on the second factor. The group $G$ is often
called the \textbf{\textit{structure group}}.
\end{definition}
%
\begin{example}[\textbf{\textit{Bundles of frames}}]
Let $M$ be a smooth manifold, and $p \in M$. Let $\mathcal{B}_p$ denote the
$GL_n\R$-torsor of frames of the tangent space $T_pM$, and let
$\mathcal{B}(M)$ be the set
\[
\mathcal{B}(M) = \coprod_{p \in M} \mathcal{B}_p
\]
This set comes with a natural map $\pi : \mathcal{B}(M) \to M$ that maps a frame
$b \in \mathcal{B}_p$ to $p$, so the fiber $\pi^{-1}(p)$ over any point is
$\mathcal{B}_p$. Local coordinates $x^i$ on an open set $U \subset M$ induce bijections
$\pi^{-1}(U) \to U \times GL_n\R$, where we map $b \in \mathcal{B}_p$ to
$(p, A_b)$, where $A_b \in GL_n\R $ is the matrix where the $i^{th}$ column are the
components of $b(e_i)$ with respect to the coordinate vectors $\partial_i$. This
then induces a topology and smooth structure on $\mathcal{B}(M)$ such that
the projection map $\pi : \mathcal{B}(M) \to M$ is smooth and the maps
$\pi^{-1}(U) \to U \times GL_n\R$ are smooth local trivializations. \\

This construction works for vector bundles as well. For a rank $k$ vector bundle
$E \to M$, there is a principal $GL_n\R$-bundle $\mathcal{B}(E) \to M$, where the fiber
over a point $p$ is the $GL_k\R$-torsor of frames for the fiber $E_x$.
\end{example}
%
Heuristically, principal bundles can be thought of bundles of symmetries of some
object, and in many cases, this object is a vector bundle, as we will soon see.
%
\section*{Groups Control the Geometry}
%
Given any geometric object $M$ (e.g. a vector space, manifold, fiber bundle), an
extremely important question to as is -- ``what are the symmetries of $M$?" More
concretely, we would like to know the \emph{automorphisms} of $M$. Any quantity
intrinsic to the object $M$ must be invariant under these automorphisms, sine
we have no ``preferred frame of reference" for the object $M$. In this way, we
see that any geometric properties of $M$ are defined by the symmetry group $G$,
and we say the group $G$ ``controls" the geometry of $M$. For example, given
a finite dimensional vector space $V$, the group of symmetries is the general
linear group $GL(V)$ of invertible linear transformations. If we further fix an
inner product $\langle\cdot,\cdot\rangle$ on $V$, we have a smaller class of
symmetries -- the orthogonal group $O(V) \subset GL(V)$ of linear automorphisms
preserving the inner product. Additional structure (e.g. orientation, symplectic
form, complex structure) allows us to pick out subgroups  of automorphisms, which gives
a more restrictive class of symmetries of our vector space, and these subgroups
define the geometry of the space. \\

This proves to be a very fruitful philosophy for approaching geometry. Smooth manifolds
are locally modeled on vector spaces, and given a structure on the linear world
of vector spaces, we often get an analogous structure in the nonlinear world of
manifolds. For example, a Riemannian metric $g$ on a manifold $M$ is the nonlinear
analogue of an inner product on a vector space. Because of this, we would expect
the orthogonal group $O_n$ to play an important role in the geometry of a Riemannaian
manifold. The interaction of the group with the geometry comes in play through the
language of principal bundles. Before we mentioned that principal bundles can be
thought of as bundles of symmetries, which we make precise with the construction of
a \textbf{\textit{associated bundle}}.
%
\begin{definition}
Let $P \to M$ be a principal $G$-bundle, and $F$ a smooth manifold with a smooth
left $G$-action. The \textbf{\textit{associated bundle}} is the bundle
\[
P \times_G F = P \times F / (p, f) \sim (p \cdot g, g^{-1} \cdot f)
\]
This is a fiber bundle over $M$ with model fiber $F$, and it is a good exercise to see
why this is true by explicitly constructing the local trivializations in terms
of local trivializations of $\mathcal{B}(M)$.
\end{definition}
%
The construction is a bit obtuse, so we make a few observations to make sense of
why we want to care about associated bundles.
%
\begin{example}[\textbf{\textit{The tangent bundle}}]
Let $M$ be a smooth manifold. From $M$, we have a principal $GL_n$-bundle
$\mathcal{B}(M) \to M$ of frames for $TM$. In addition, $GL_n\R$ admits a natural
left action on $\R^n$ via matrix multiplication. The associated bundle
$E = \mathcal{B}(M) \times_{GL_n\R} \R^n$ is isomorphic to the tangent bundle
$TM$. To see this, we construct maps in both directions. \\

Let $[b, v]$ denote an equivalence class in $E$, where $b \in \mathcal{B}_p$ is a
frame for $T_pM$ and $v \in \R^n$. Then define $\varphi : E \to TM$ by
$\varphi[b,v] = b(v)$. This is well defined, since
\[
\varphi[b\cdot g, g^{-1}\cdot v] = b \circ g(g^{-1}(v)) = b(v)
\]
In the other direction, let $(p,v) \in TM$, i.e. $p \in M$ and $v \in T_pM$. Fix
a basis $b : \R^n \to T_pM$, and define $\psi : TM \to E$ by
$\psi(p,v) = [b,b^{-1}(v)]$. It's an easy exercise to check that these two maps compose
to identity in both directions.
\end{example}
%
The above construction works in the setting of vector bundles as well. Given
a rank $k$ vector bundle $E \to M$, we can construct the principal $GL_k\R$-bundle
$\mathcal{B}(E)$, and the associated bundle $\mathcal{B}(E) \times_{GL_k\R} \R^k$ is
isomorphic to the bundle $E$, which tells us that the processes of taking frame bundles
and constructing associated bundles are inverses to each other -- no information
is lost in either direction, so we can work in whichever setting proves most
convenient. \\
%

This construction also shows why we might care about associated bundles -- the quotient
by relation we specified exactly gives the correct transformation law for tangent
vectors. You may have heard the joke that physicists define a vector as ``something
that transforms like a vector." A more precise joke here would perhaps be ``a tangent
vector is something that transforms like a tangent vector." What do we mean by this?
Physically, the only ``real" quantities are those invariant under a change of reference
frame. In terms our example, we see that a tangent vector $v \in T_pM$ is
\emph{not} just an $n$-tuple of numbers, it's the collection of all coordinate
representations of $v$ with respect to any basis of $T_pM$, which is exactly what
the associated bundle construction captures. We can see this another way. Fix
an element $b \in \mathcal{B}_p \subset \mathcal{B}(M)$. This then determines an
isomorphism of $\R^n$ to the fiber of the associated bundle
$\mathcal{B}(M) \times_{GL_n\R} \R^n$ by $v \mapsto [b,v]$, this exactly what happens
when we fix a basis for $T_pM$! In this way, we see that fixing the first component
in the equivalence class of an element of an associated bundle is essentially a
choice of basis or reference frame, and doing the computations with the equivalence
classes themselves is in essense, working in a coordinate-free manner by working
in all coordinate systems simultaneously.\\

%TODO sections are equivalent to G-equivariant maps
Another important property of the associated bundle construction is that we can
interpret sections of any associated bundle in terms of maps on the principal bundle.
%
\begin{proposition}
Let $P \to M$ be a principal $G$-bundle, and $F$ a manifold with a left $G$-action, and
let $E = P \times_G F$ denote the corresponding associated bundle. Then there
is a bijective correspondence
\[
\Gamma_M(E) \longleftrightarrow \{G\text{-equivariant maps } P \to F\}
\]
\end{proposition}
%
\begin{proof}
We provide maps in both directions. First, let $\psi : M \to E$ be a section. We
then define the corresponding $G$-equivariant map $\widetilde{\psi} : P \to F$.
We know that for any $x \in M$, we have that $\psi(x) = [p, f]$ for $p \in P_x$.
Then define $\widetilde{\psi}(p)$ to be the second factor of the unique representative
of $[p,f]$ with $p$ as a first factor. By the definition of the associated bundle,
this map is $G$-equivariant. \\

In the other direction, let $\widetilde{\varphi} : P \to F$ be a $G$-equivariant map.
We then construct its corresponding section $\varphi : M \to E$. For $x \in M$,
let $\varphi(x) = [p,\tilde{\varphi}(p)]$ for any $p \in P_x$. The definition of
the associated bundle ensures that this is well defined, and it is easy to see that
the two maps we defined are inverses to each other.
\end{proof}
%TODO comment on working on bundle or on the base
This is a specific example of a very general phenomenon -- we can either work on the
base manifold $M$, or we can work $G$-equivariantly on the principal bundle $P \to M$
and use the $G$-equivariance to descend to the base manifold. The idea of working
$G$-equivariantly captures the idea in all coordinate systems at once. \\
%

We now address how fixing additional structures on our manifold $M$ changes this
picture. The best example here is that of a Riemannian metric $g$ on $M$, which gives
us a notion of an \textbf{\textit{orthonormal frame}} of the tangent space $T_pM$,
which is a linear \emph{isometry} $b : (\R^n,\langle\cdot,\cdot\rangle) \to (T_pM,g_p)$
where $\langle\cdot,\cdot\rangle$ is the standard inner product on $\R^n$, and
$g_p$ is the Riemannian metric evaluated at the point $p$. We then construct the
\textbf{\textit{orthonormal frame bundle}} $\mathcal{B}_O(M)$ in the same way we
constructed the frame bundle $\mathcal{B}(M)$, which is a principal $O_n$-bundle,
which reflects that we now have a more restrictive view on symmmetry -- our
automorphisms now take values in $O_n$ instead of $GL_n\R$.
%
\begin{definition}
Let $M$ be a smooth manifold, and $\mathcal{B}(M)$ its $GL_n\R$-bundle of frames.
Let $G$ be a Lie group, and $\rho : G \to GL_n\R$ be a homomorphism. Then a
\textbf{\textit{reduction of structure group}} to $G$ is the data of a principal
$G$-bundle $Q \to M$ equipped with a $G$-equivariant map $Q \to \mathcal{B}(M)$, where
$G$ acts on the left of $\mathcal{B}(M)$ by $g \cdot b = b \circ \rho(g)^{-1}$.
\end{definition}
%
In the case we illustrated above with a Riemannian metric $g$, the homomorphism
$\rho$ is just the inclusion $O_n \hookrightarrow GL_n\R$, the bundle $Q$ is
$\mathcal{B}_O(M)$, and the $O_n$-equivariant map is the inclusion
$\mathcal{B}_O(M) \hookrightarrow \mathcal{B}(M)$. However, the homomorphism $\rho$
does \emph{not} need to be injective, which makes the name a bit a misnomer -- the
group does not need to be a subgroup of $GL_n\R$. For example, when working with
the Spin group $\mathrm{Spin}_n$, the map $\rho$ is often the double cover
$\mathrm{Spin}_n \to SO_n$. \\

Our construction of the tangent bundle as an associated bundle can also be generalized
to other tensor bundles built out of the tangent bundle (e.g. $\mathcal{T}^k_\ell(M)$,
$T^*M$, $\Lambda^k(T^*M)$, etc), by using the induced representations of the
stucture group $G$ on tensor products, the dual representation, and the induced
representations on exterior powers respectively. From this, we see that the associated
bundle construction is the bridge between the theory of principal bundles and vector
bundles with representation theory, and reductions of structure group allows us to
import our linear model geometries (e.g. an inner product space, a vector space with
a complex structure $I$, an oriented vector space, etc.) to the nonlinear world of
manifolds by asking for a reduction of structure group to the appropriate group. For
example, an orientation is equivalent to a reduction of structure group to
$GL_n^+\R = \{A \in GL_n\R ~:~ \det A > 0\}$, and an almost complex structure on a
$2n$-dimensional manifold is equivalent  to a reduction of structure group from
$GL_{2n}\R$ to $GL_n\C$.
%
\section*{Connections}
%
Another vital piece of the theory of principal bundles is the theory of connections.
Unfortunately, this discussion  will require some additional background. This section
will assome some knowledge of algebraic topology (on the level over covering spaces),
and a decent amount of comfort with differential geometry. \\

We first motivate why we would care about connections. Suppose our group $G$ is finite.
Then a principal $G$-bundle $P \to M$ is the same as a (possibly disconnected)
$|G|$-fold cover of $M$ equipped with a free and transitive $G$ action. For example,
double covers of $M$ are exactly principal $\Z/2\Z$-bundles over $M$. An extremely
important property of covering spaces is the \textbf{\textit{path lifting property}}.
Given a cover $P \to M$, a point $x \in M$, and a lift $p \in P_x$ of $x$, let
$\gamma : I \to M$ be a path with $\gamma(0) = x$. Then there exists a unique
path $\widetilde{\gamma} : I \to P$ such that $\widetilde{\gamma}(0) = p$, and
$\pi \circ \widetilde{\gamma} = \gamma$. We call the path $\widetilde{\gamma}$ a
\textbf{\textit{lift}} of $\gamma$. This makes use of the fact that there is a
neighborhood about $p$ that is mapped diffeomorphically onto a neighborhood of $x$. \\

We would also like a path lifting property for a general principal bundle
$P \to M$, but our hopes are immediately dashed by a simple dimension count --
the dimension of $P$ is $\dim G + \dim M$, so if $G$ is not discrete, we have
no hope of lifting a path, since there will be no local diffeomorphisms we will be
able to use to lift a path. To accomplish our goal of path lifting, we need
additional data.
%
\begin{definition}
Let $M$ a smooth manifold. A \textbf{\textit{rank $k$ distribution}} $H$ over $M$ is
a vector subbundle $H \subset TM$, such that the fiber $H_x$ over any $x \in M$
is a $k$-dimensional subspace of $T_xM$.
\end{definition}
%
A distribution can be thought of as a higher dimensional analogue of a vector field,
where instead of tangent vectors, we are asking for tangent $k$-planes. A
principal bundle comes equipped with a natural distribution.
%
\begin{definition}
Let $\pi : E \to M$ be a fiber bundle with model fiber $F$.
Then the subbundle $\ker\pi_*$ is a distribution of rank $\dim F$ called the
\textbf{\textit{vertical distribution}}.
\end{definition}
%
Since a fiber bundle is locally modeled on a product $U \times F$, we can
think of it locally as several copies of $F$ sitting over out manifold $M$. The
vertical distribution consists of all the directions in the fiber. If you move in
any of the directions of the vertical distribution, you stay in the same fiber
you started with.

In the special case that we are working with a principal bundle, an important fact is
that the vertical distribution has a canonical identification with the product bundle
$M \times \mathfrak{g}$ of $M$ with the Lie algebra $\mathfrak{g}$ of $G$.
%
\begin{proposition}
Let $P \to M$ be a principal $G$-bundle with vertical distribution $V$. Then
every fiber $V_p$ of the vertical distribution has a canonical identification with
$\mathfrak{g}$.
\end{proposition}
%
\begin{proof}
Using the identification of $\mathfrak{g}$ with left-invariant vector fields on
$G$, there is an \textbf{\textit{exponential map}} $\exp : \mathfrak{g} \to G$
where for an element $X \in \mathfrak{g}$, we define $\exp(X) = \gamma_X(1)$, where
$\gamma_X$ is the integral curve of the vector field $X$ with $\gamma_X(0) = e$. The
exponential map maps a small line segment
$\{tX ~:~ t \in (-\varepsilon, \varepsilon)\}$ to a curve
$\gamma : (-\varepsilon,\varepsilon) \to G$, with $\gamma(0) = e$. Fix a point $p \in P$.
Then any Lie algebra element $X \in \mathfrak{g}$ determines a tangent vector
$\widehat{X} \in T_pP$ by the formula
\[
\widehat{X} = \frac{d}{dt}\bigg\vert_{t = 0}p \cdot \exp(tX)
\]
Since the group action preserves the fibers, $\widehat{X} \in V_p$, and it is a
useful exercise to check that the mapping $X \mapsto \widehat{X}$ is an isomorphism
$\mathfrak{g} \to V_p$.
\end{proof}
%
There is an analogous identification for a vector bundle.
%
\begin{proposition}
Let $\pi : E \to M$ be a vector bundle with vertical distribution $V$. Then for
$e \in E$, there is a canonical identification of $V_e$ with $E_{\pi(e)}$, where
the mapping is given by
\[
v \mapsto \frac{d}{dt}\bigg\vert_{t=0}e + tv
\]
\end{proposition}
%
\begin{definition}
Let $\pi : E \to M$ be a fiber bundle with vertical distribution $V$.
Then a \textbf{\textit{connection}} is a choice of
\textbf{\textit{horizontal distribution}}, i.e. a distribution $H \subset TE$ such that
for every $p \in P$, we have that $V_p \oplus H_p = T_pE$.
\end{definition}
%
A connection on $E$ gives us distinguished ``manifold directions" which are
complementary to the ``fiber directions" defined by the vertical distribution $V$.
By the first isomorphism theorem, the differential
$\pi_*\vert_{H_p} : H_p \to T_{\pi(p)}M$ is an isomorphism, which is the
essential fact we will use to define path lifting on a fiber bundle. \\

Specializing to principal bundles $P \to M$, one
useful thing to note is that the choice of complementary subspace to $V_p$
is equivalent data to the choice of a projection map $T_pP \to V_p$, with $H_p$ as
the kernel. Using the identification of $V_p$ with $\mathfrak{g}$, this is
equivalent to a surjective map $T_pP \to \mathfrak{g}$. Doing this fiberwise, we
see that a connection $H$ is equivalent data to a $\mathfrak{g}$ valued $1$-form
$\Theta \in \Omega^1_P(\mathfrak{g})$, i.e. a section of
$\Lambda^1(P) \otimes \mathfrak{g}$, where we recover $H$ as the kernel of $\Theta$.
In addition, since the projection map is identity on $V$, it is clear that
for $X \in \mathfrak{g}$, we have that $\Theta(\widehat{X}) = X$. The form
$\Theta$ is called the \textbf{\textit{connection form}}. An analogous construction
can be made for a vector bundle $\pi: E \to M$ as well, giving a connection
form $\omega \in \Omega_E^1(\pi^*E)$ valued in the pullback bundle $\pi^*E$. \\
%

We now return to the issue of path lifting. Given a fiber bundle $\pi : E \to M$
equipped with a connction $H \subset TE$, we remarked earlier that
$\pi_*\vert_{H_p} : H_p \to T_{\pi(p)}M$ is an isomorphism. This means that for
any tangent vector $v \in T_{\pi(p)}$, there is a \emph{unique} vector
$\tilde{v} \in H_p \subset T_pE$ such that $\pi_*(\tilde{v}) = v$. We call
$\tilde{v}$ the \textbf{\textit{horizontal lift}} of the vector $v$. We
now have all the ingredients in place for path lifting. \\

Let $\pi : E \to M$ be a fiber bundle equipped with connection $H$, and let
$\gamma : I \to M$ be a smooth path with $\gamma(0) = x$. For each time $t \in I$,
we have a velocity vector $\gamma'(t) \in T_{\gamma(t)}M$. This determines a vector
field $V$ on the fibers along the path $\gamma$, where the value of $V$ at any point
$p \in P_{\gamma{t}}$ is the unique horizontal lift of $\gamma'(t)$ to $p$. By
the theory of differential equations, we can integrate these vector fields to a flow
on the fibers along the path, so fixing a point $p \in P_x$ determines a
unique integral curve $\widetilde{\gamma} : I \to P$ starting at $p$, which satisfies
$\pi \circ \widetilde{\gamma} = \gamma$, solving the path lifting problem. We
can then use the lifted paths to define \textbf{\textit{parallel transport}} along
a curve. Given a curve $\gamma I \to M$, we get a family of maps
$\tau_t : E_{\gamma(0)} \to E_{\gamma}(t)$, where $\tau_t(f) = \gamma_f(t)$, where
$\gamma_f$ is the unique lift of $\gamma$ to a path starting at $f \in E_{\gamma(0)}$.
These maps are isomorphisms, since they have an inverses given by the parallel transport
maps induced by the path $\gamma$ traversed backwards. The parallel transport
maps give us a way to transport data between fibers along a curve, so a connection
``connects" nearby fibers, justifying the name. Our usage of horizontal lifts of
velocity vectors shows us why the connection was the missing data for path lifting.
Since our group was not discrete, we did not have a distinguished ``manifold" direction
to lift the path, which was exactly the problem that the horizontal distribution solved.
In this way, we see that a recipe for path lifting is equivalent data for a connection,
since any recipe for path lifting needs to specify a distinguished ``manifold" direction
to connect nearby fibers.

We now restrict our attention to principal bundles. An extremely important fact
is that a connection on a principal $G$-bundle $P \to M$ induces a connection
on \emph{any} associated bundle $P \times_G F$. There are several ways to see this.
One way is through the horizontal distribution. Let $H \subset TP$ be the horizontal
distribution on $P$. Then $T(P \times F)$ is canonically isomorphic to $TP \oplus TF$,
so we get a distribution $H \times \{0\} \subset TP \oplus TF$. This is invariant
under the $G$ action on $P \times F$, so it descends to the quotient $P \times_G F$,
giving us a horizontal distribution on the associated bundle. Another way to
see this is that the connection on $P$ gives us path lifting on any associated bundle.
Fix a basepoint $x \in M$, and a basepoint $(p, f)$ in the fiber of $P \times_G F$
over $x$. Then given a path $\gamma : I \to M$ based at $x$, we get a unique lift
$\widetilde{\gamma} : I \to P$ to a curve in $p$. We then get a path
$\widehat{\gamma} : I \to P \times_G F$ defined by
$\widehat{\gamma}(t) = [\widetilde{\gamma}(t), f]$. This gives a useful intepretation
of curves on fiber bundles with horizontal velocity vectors. Thinking of the
principal bundle component as a sort of ``coordinate system," These curves
are constant in the second component, and the connection on $P$ allows us to pick a
distinguished path of frames. In the case of the tangent bundle, we can say
even more. Given a path $\gamma : I \to M$ in $M$, the mapping
$t \mapsto (\gamma(t), \gamma'(t))$ defines a curve in $TM$. Thinking of $TM$ as the
associated bundle $\mathcal{B}(M) \times_{GL_n\R} \R^n$, what does it mean
for this curve in $TM$ to be of the form $[\widetilde{\gamma}(t), v]$ for a fixed
$v \in \R^n$? We can think of $v$ as the \emph{acceleration}. Therefore, asking for
the curve in $TM$ to be horizontal is like asking for the curve $\gamma$ on $M$ to
be \emph{constant accleration}, which is one definition of a geodesic.
%TODO Covariant derivative.
\section*{References}
%
Most of the content of this article came from many discussions with Professor Dan
Freed. If you want to learn more, I heavily recommend finding some time to talk to
him.  A good reference for the theory of principal bundles can be found in Kobayashi
and Nomizu's books \emph{Foundations of Differential Geometry}.
%
\end{document}